% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{Conclusion}\label{chapter:conclusion}

This work surveyed numerous offline and online algorithms for fractional and integral smoothed convex optimization in single and high dimensions. We evaluated the performance of online algorithms for the application of dynamically right-sizing data centers using traces from various different sources. We found that the discussed online algorithms perform nearly optimally with respect to the observed normalized cost when compared to the dynamic offline optimum. Moreover, we observed that the normalized cost increases as potential cost savings increase. However, the achieved cost reduction depends primarily on the difference between the dynamic offline optimum and the strategy compared to (in our experiments the static offline optimum). Previous results show that changes in the data center model, such as increasing the energy cost or decreasing the normalized switching cost lead to significant increases in cost savings. We also observed that the difference between fractional and integral solutions is negligible in the application of right-sizing data centers, or generally when total costs are large. Further, we have seen that online algorithms using predictions of future loads can achieve a better normalized cost, even when predictions are non-perfect but obtained using standard methods. However, we have also seen that the achieved normalized cost does not necessarily improve for larger prediction windows.

Finally, we made our implementation of the discussed algorithms available. This implementation can be used to test the performance of algorithms for smoothed online convex optimization in other applications and compare their performance. Our data center model balances energy cost and revenue loss. Using past data, this model can be tuned to be safe (i.e., provide enough computing resources for incoming loads), fulfill service level agreements, and provide significant cost and energy savings in real-world settings.